---
title: ""
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```
<p>&nbsp;</p>

# About <tt>`R`</tt> and <tt>`RStudio`</tt>
R is a free open-source programming language that can be used widely for statistical analysis. Another free program is Rstudio, that provides a nice graphic interface for R. We need to download R first, then download Rstudio. Both can run on PCs, Macs or Linux.

# Introduction to <tt>`R`</tt>
This is a mini course about R programming language with some useful and important functions that we need to know before we learn more advanced and tricky functions. I hope you enjoy it and please feel free to ask questions. 

We can use R as a calculator. We can use addition, subtraction, multiplication and division very easily in R. For example,

```{r,comment=NA}
15+20
12*12 # * is multiplication sign.
126/3 # / is division sign.
(2+5+7)/5
```
if you want to know more information about any function in R, you need to type help("function name") or ?function name.

- Notice that everyting comes after hashtag \# will be ignored by R.

<!-- # Laplace approximation -->
<!-- If we have one-dimensional numerical integration, to find out that integration. We could possibly use a Laplace approximation. -->


<!-- # Example from my thesis. -->

# Variable assignment

Put numbers into variables, do simple math on the variables

```{r, comment=NA}
a<-1
b<-2
a+b

d<-c(1,2,3)
e<-c(5,6,7)
d+e
d*e
d/e

```

# Data types
There are different types of data in R such as double, character and logical data.

- Double data: When we deal with "numbers" in R, then the type of data is called <span style="color:blue;">double</span>.

We can check if the variable is double or not using the following function.
```{r,comment=NA}
x<-5
is.double(x)
```
So the answer is true. That means the variable x is double in R.


- Character data:

- Logical data:

# Vectors

# Matrices



# Simple stuff to do with some <tt>`R`</tt>

If you would like to make a histogram in R. For instance, we want to generate a sample with 100 numbers from a normal distribution with mean = 0, and standard deviation = 0.5, and then we want to plot a histogram. We can do this here by using the folowing R code:

```{r}
samples <- rnorm(100, mean=0, sd=0.5)
hist(samples,breaks=10)
#hist(samples, prob)
#lines(density(samples),col=4)
```


# What can you do right now by yourself?

It's hard to learn programming when you don't have specific problems that you are trying to solve. Everything just seems abstract.

I wrote an [introductory programming book that introduces R](https://crumplab.github.io/programmingforpsych/), and gives some [concrete problems for you to solve](https://crumplab.github.io/programmingforpsych/programming-challenges-i-learning-the-fundamentals.html). 


# Problem 3

Write code that will place the numbers 1 to 100 separately into a variable using for loop. Then, again using the seq function.

```{r, comment=NA}
# for loop solution
# i becomes the number 1 to 100 at each step of the loop


a <- length(20) # make empty variable, set length to 100
for (i in 1:20){
  a[i] <-i #assigns the number in i, to the ith index of a
}

print(a)

# for loop solution #2

a<-c() #create empty variable using combine command
for (i in 1:20){
  a<-c(a,i) # keeps combining a with itself and the new number in i
}
print(a)

# seq solution
a <- seq(1,20,1) # look up help for seq using ?seq() in console
print(a)

```

# Replace this with problem 4

And keep going. Try to solve the problems with different scripts that provide the same solution. 

# Problem 5
```{r,comment=NA}
z<-rnorm(10,0,0.1)
hist(z,probability=TRUE)
lines(density(z),col=4)
(z+5)
summary(z) ### Give summary of the distribution
```

# Laplace approximation
```{r,comment=NA}
### It is true and I checked it for Bernoulli likelihood and
### normal prior distribution with variance=1. Albumin has 0 and 1.

Laplacealbumin<-function(y,E0,tol=1E-5,n=20)
{#MF 04/10/18
  z0<-E0
  d<-1
  for (i in 1:n)
  {if (abs(d)>tol)
  {ez<-exp(z0)
  pz<-ez/(1+ez)
  f<-y-(z0-E0)-pz
  fd<--(1+pz*(1-pz))
  d<-f/fd
  z0<-z0-d
  }
  }
  mean<-z0
  var<- -1/fd
  return(list(E1=mean,V1=var))
}

Laplacealbumin(y=1,E0=0,tol=1E-5,n=20)
```

Simple function to find the mean and the variance of a set of data.
```{r,comment=NA}
## Simple function to find the mean and the variance for 
## set of data.

f<-function(x,n)
{
  xbar<-mean(x)
  n<-length(x)
  ss<-sum((x-xbar)^2)
  v<-ss/(n-1)
  return(list(xbar=xbar,v=v))
}

x<-c(1:3)
f(x,n)
```
Now to check if our function is correct, we have intrinsic functions or built-in functions in R to calculate the mean and the variance for data set. So,
```{r, comment=NA}
mean(x);var(x)

```


Good luck and Happy coding. Thanks

